Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2020-01-31 doc/Paper/main.tex
Small changes.

  \subsubsection{Photometric redshift systematics modeling}\label{sssec:methods.theory.photoz_syst}
    Uncertainties in photometric redshifts represent one of the major systematic uncertainties in photometric galaxy clustering analyses. \rev{In this work, we choose to model photo-$z$ errors as causing shifts in the means and changes in the width of the derived redshift distribution for a population of galaxies\footnote{We note our photo-$z$ error model does not take into account the full effect of catastrophic outliers but we do test for the presence of this systematic in Sec.~\ref{sssec:results.pz}.}. This allows us to parametrize the derived galaxy redshift distributions $p_{i}(z)$ using a two-parameter model given by}
    \begin{equation}
      p_{i}(z) \propto \hat{p}_{i}(z_{c} + (1 + z_{w, i})(z-z_{c}) + \Delta z_{i}),
      \label{eq:photo-z-model}
    \end{equation} 
    where the index $i$ runs over the number of redshift bins considered in our analysis. In the above equation, $\hat{p}_{i}(z)$ denotes the estimated redshift distribution while $p_{i}(z)$ is the underlying true distribution. The parameters $\Delta z_{i}$ account for shifts in the means of the distributions and changes to their widths are parameterized through $z_{w, i}$. The quantity $z_{c}$ is kept constant in our analysis and is set to the redshift at which $\hat{p}_{i}(z)$ attains its maximal value. The normalization of the redshift distribution $p_{i}(z)$ depends on both $\Delta z_{i}$ and $z_{w, i}$ and we do not account for it in Eq.~\ref{eq:photo-z-model} for clarity. \rev{It is worth noting that we do not expect for this parametrization to fully capture the complications of realistic catastrophic outliers, and we quantify its ability to account for the underlying photo-$z$ uncertainty given our statistical errors in Section \ref{sssec:results.pz}.}

@@

    Finally, we compute the super-sample covariance contribution following the treatment of \cite{Krause:2017}, i.e.:
     \begin{equation}
    \begin{aligned}
       \mathrm{Cov}_{\mathrm{SSC}}(C^{ab}_{\ell}, C^{cd}_{\ell'}) &= \int \mathrm{d}\chi \;\frac{q^{a}(\chi)q^{b}(\chi)q^{c}(\chi)q^{d}(\chi)}{\chi^{4}} \times \\ &\frac{\partial P_{ab}(\sfrac{\ell}{\chi}, z(\chi))}{\partial \delta_{\rm LS}}\frac{\partial P_{cd}(\sfrac{\ell'}{\chi}, z(\chi))}{\partial \delta_{\rm LS}}\sigma^{2}_{b}(z(\chi)).
    \end{aligned}
    \end{equation}
        The quantity $\sigma_b^2(z)$ is the variance of the long wavelength mode $\delta_{\rm LS}$ over the survey footprint, given by
    \begin{equation}
      \sigma_b^2(z) = \int \frac{\mathrm{d}k_\perp^2}{(2\pi)^2}P_{\rm lin}(k_\perp,z)\left|W(k_\perp,z)\right|^2.
    \end{equation}
    Furthermore $W(k_\perp,z)$ denotes the Fourier transform of the survey footprint, which we approximate as a compact circle with an area matched to our data set\footnote{\rev{We expect this approximation to be acceptable for the present analysis as HSC covers a small sky fraction and we mainly consider large angular multipoles (see e.g. Refs.~\cite{Krause:2017a, Barreira:2018}).}}:
    \begin{equation}
      W(k_\perp,z)=\frac{2 J_1(k_\perp\chi(z)\theta_s)}{k_\perp \chi(z)\theta_s},\hspace{12pt} \theta_s={\rm arccos}(1-2f_{\rm sky}),
    \end{equation}
    where $J_1(x)$ is the cylindrical Bessel function of order 1. \rev{It is worth noting that the impact of the SSC on the covariance matrix for galaxy clustering is reduced due to the integral constraint, and therefore this approximation does not affect our results significantly (see also Sections \ref{ssec:results.covariance} and \ref{sssec:results.choices}).}
    The quantity $\partial P_{ab}(k, z)/\partial \delta_{\rm LS}$ is the response of the power spectrum $P_{ab}$ to a large-scale density fluctuation, which we estimate using the halo model and results from perturbation theory as (e.g. \cite{Krause:2017}):
     \begin{equation}
    \begin{aligned}
      \frac{\partial P_{ab}(k, z)}{\partial \delta_{\rm LS}} =& \left( \frac{68}{21} - \frac{1}{3}\frac{\mathrm{d}\log{k^{3} P_{\mathrm{lin}}}(k, z)}{\mathrm{d}\log k} \right) I_{a}^{1}(k)I_{b}^{1}(k)P_{\mathrm{lin}}(k, z) + I_{ab}^{1}(k, k) \\ &- (b_{a} + b_{b})P_{ab}(k, z).
    \label{eq:ps-resp}  
    \end{aligned}
     \end{equation}
    The last term in Eq.~\ref{eq:ps-resp} accounts for the fact that the observed galaxy overdensity is computed using the mean galaxy density estimated inside the survey volume.

@@

    \begin{figure}
      \begin{center}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=0.pdf}
        \caption{$C_{\ell}^{11}$}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=1.pdf} \\
        \caption{$C_{\ell}^{22}$}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=2.pdf}
        \caption{$C_{\ell}^{33}$}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=3.pdf}
        \caption{$C_{\ell}^{44}$}
        \end{subfigure}
        \caption{Comparison of the different contributions to the covariance matrix for the four auto-power spectra considered in our analysis.} \label{fig:covariance-contributions}
      \end{center}
    \end{figure}

@@

  Despite the small fraction of sky covered by the data used in this work, HSC DR1 constitutes an ideal proxy for the data expected for LSST. Our analysis demonstrates the feasibility of photometric galaxy clustering studies using such high signal-to-noise datasets. In particular, this work shows the level of readiness of the LSST DESC analysis pipelines, and showcases the application of power spectrum methods for galaxy clustering, and the use of template deprojection techniques to handle contamination from sky systematics. We anticipate that future development of this pipeline will include improvements to photo-$z$ marginalization and non-linear bias modeling, thus permitting its application to magnitude-limited galaxy samples with even higher signal-to-noise. \rev{In particular, the increased statistical power of LSST will require not only a more rigorous marginalization over photo-$z$ uncertainties, but also a calibration sample significantly larger than the COSMOS 30-band catalog use here in order to reduce the associated sample variance.} These results therefore bode well for future cosmological constraints using galaxy clustering data from LSST.

@@

  The contributions from the authors are listed below:
  AN: co-led project, led signal modeling, photo-$z$ systematics characterization, likelihood analysis and interpretation of results.
  DA: co-led project, led development of the data analysis pipeline, contributed to likelihood and modelling efforts.
  JS: development of data analysis pipeline, comparisons with simulated data, checked impact of observing conditions (systematics).
  AS: co-led project, advised student, cross checked MCMC chains, calculated $N(z)$ sample variance uncertainty, contributed to modelling efforts, wrote initial likelihood code.
  HA: carried out initial data exploration, analyzed HSC photo-$z$ pdfs, and determined tomographic bin selection.
  AB: shot noise estimation, small updates to HSC pipeline.
  JD: provided advice and guidance.
  ZG: calibrated redshift distributions from COSMOS 30-band data.
  EG: galaxy sample selection and shot noise estimation.
  RM: provided guidance on how to use the HSC data.
  HM: provided guidance on how to use the HSC data.
  JN: assistance in interpretation of galaxy clustering results.
  SS: model parametrization, testing MCMC chains, studied uncertainties in the first bin's $N(z)$.
  EW: produced observing condition maps.

>>>>

  \subsubsection{Photometric redshift systematics modeling}\label{sssec:methods.theory.photoz_syst}
    Uncertainties in photometric redshifts represent one of the major systematic uncertainties in photometric galaxy clustering analyses. \rev{In this work, we choose to model photo-$z$ errors as causing shifts in the means and changes in the width of the derived redshift distribution for a population of galaxies. This allows us to parametrize the derived galaxy redshift distributions $p_{i}(z)$ using a two-parameter model given by}
    \begin{equation}
      p_{i}(z) \propto \hat{p}_{i}(z_{c} + (1 + z_{w, i})(z-z_{c}) + \Delta z_{i}),
      \label{eq:photo-z-model}
    \end{equation} 
    where the index $i$ runs over the number of redshift bins considered in our analysis. In the above equation, $\hat{p}_{i}(z)$ denotes the estimated redshift distribution while $p_{i}(z)$ is the underlying true distribution. The parameters $\Delta z_{i}$ account for shifts in the means of the distributions and changes to their widths are parameterized through $z_{w, i}$. The quantity $z_{c}$ is kept constant in our analysis and is set to the redshift at which $\hat{p}_{i}(z)$ attains its maximal value. The normalization of the redshift distribution $p_{i}(z)$ depends on both $\Delta z_{i}$ and $z_{w, i}$ and we do not account for it in Eq.~\ref{eq:photo-z-model} for clarity. \rev{It is worth noting that we do not expect for this parametrization to fully capture the complications of realistic catastrophic outliers, and we quantify its ability to account for the underlying photo-$z$ uncertainty given our statistical errors in Section \ref{sssec:results.pz}.}

@@

    Finally, we compute the super-sample covariance contribution following the treatment of \cite{Krause:2017}, i.e.:
     \begin{equation}
    \begin{aligned}
       \mathrm{Cov}_{\mathrm{SSC}}(C^{ab}_{\ell}, C^{cd}_{\ell'}) &= \int \mathrm{d}\chi \;\frac{q^{a}(\chi)q^{b}(\chi)q^{c}(\chi)q^{d}(\chi)}{\chi^{4}} \times \\ &\frac{\partial P_{ab}(\sfrac{\ell}{\chi}, z(\chi))}{\partial \delta_{\rm LS}}\frac{\partial P_{cd}(\sfrac{\ell'}{\chi}, z(\chi))}{\partial \delta_{\rm LS}}\sigma^{2}_{b}(z(\chi)).
    \end{aligned}
    \end{equation}
        The quantity $\sigma_b^2(z)$ is the variance of the long wavelength mode $\delta_{\rm LS}$ over the survey footprint, given by
    \begin{equation}
      \sigma_b^2(z) = \int \frac{\mathrm{d}k_\perp^2}{(2\pi)^2}P_{\rm lin}(k_\perp,z)\left|W(k_\perp,z)\right|^2.
    \end{equation}
    Furthermore $W(k_\perp,z)$ denotes the Fourier transform of the survey footprint, which we approximate as a compact circle with an area matched to our data set\footnote{\rev{We expect this approximation to be acceptable for the present analysis as HSC covers a small sky fraction and we mainly consider large angular multipoles (see e.g. Refs.~\cite{Krause:2017a, Barreira:2018}).}}:
    \begin{equation}
      W(k_\perp,z)=\frac{2 J_1(k_\perp\chi(z)\theta_s)}{k_\perp \chi(z)\theta_s},\hspace{12pt} \theta_s={\rm arccos}(1-2f_{\rm sky}),
    \end{equation}
    where $J_1(x)$ is the cylindrical Bessel function of order 1. \rev{It is worth noting that the impact of the SSC on the covariance matrix for galaxy clustering is reduced due to the integral constraint (c.f. Eq.~\ref{eq:ps-resp}), and therefore this approximation does not affect our results significantly (see also Sections \ref{ssec:results.covariance} and \ref{sssec:results.choices}).}
    The quantity $\partial P_{ab}(k, z)/\partial \delta_{\rm LS}$ is the response of the power spectrum $P_{ab}$ to a large-scale density fluctuation, which we estimate using the halo model and results from perturbation theory as (e.g. \cite{Krause:2017}):
     \begin{equation}
    \begin{aligned}
      \frac{\partial P_{ab}(k, z)}{\partial \delta_{\rm LS}} =& \left( \frac{68}{21} - \frac{1}{3}\frac{\mathrm{d}\log{k^{3} P_{\mathrm{lin}}}(k, z)}{\mathrm{d}\log k} \right) I_{a}^{1}(k)I_{b}^{1}(k)P_{\mathrm{lin}}(k, z) + I_{ab}^{1}(k, k) \\ &- (b_{a} + b_{b})P_{ab}(k, z).
    \label{eq:ps-resp}  
    \end{aligned}
     \end{equation}
    The last term in Eq.~\ref{eq:ps-resp} accounts for the fact that the observed galaxy overdensity is computed using the mean galaxy density estimated inside the survey volume.

@@

    \begin{figure}
      \begin{center}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=0.pdf}
        \caption{$C_{\ell}^{11}$}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=1.pdf} 
        \caption{$C_{\ell}^{22}$}
        \end{subfigure}\\
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=2.pdf}
        \caption{$C_{\ell}^{33}$}
        \end{subfigure}
        \begin{subfigure}{0.49\textwidth}
        \includegraphics[width=\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_comparison=G+NG+SSC_bin=3.pdf}
        \caption{$C_{\ell}^{44}$}
        \end{subfigure}
        \caption{Comparison of the different contributions to the covariance matrix for the four auto-power spectra considered in our analysis.} \label{fig:covariance-contributions}
      \end{center}
    \end{figure}

@@

  Despite the small fraction of sky covered by the data used in this work, HSC DR1 constitutes an ideal proxy for the data expected for LSST. Our analysis demonstrates the feasibility of photometric galaxy clustering studies using such high signal-to-noise datasets. In particular, this work shows the level of readiness of the LSST DESC analysis pipelines, and showcases the application of power spectrum methods for galaxy clustering, and the use of template deprojection techniques to handle contamination from sky systematics. We anticipate that future development of this pipeline will include improvements to photo-$z$ marginalization and non-linear bias modeling, thus permitting its application to magnitude-limited galaxy samples with even higher signal-to-noise. \rev{In particular, the increased statistical power of LSST will require not only a more rigorous marginalization over photo-$z$ uncertainties, but also a calibration sample significantly larger than the COSMOS 30-band catalog used here in order to reduce the associated sample variance. The results presented in this analysis} bode well for future cosmological constraints using galaxy clustering data from LSST.

@@

  The contributions from the authors are listed below:
  AN: co-led project, led signal modeling, photo-$z$ systematics characterization, likelihood analysis and interpretation of results.
  DA: co-led project, led development of the data analysis pipeline, contributed to likelihood and modeling efforts.
  JS: development of data analysis pipeline, comparisons with simulated data, checked impact of observing conditions (systematics).
  AS: co-led project, advised student, cross checked MCMC chains, calculated $N(z)$ sample variance uncertainty, contributed to modeling efforts, wrote initial likelihood code.
  HA: carried out initial data exploration, analyzed HSC photo-$z$ pdfs, and determined tomographic bin selection.
  AB: shot noise estimation, small updates to HSC pipeline.
  JD: provided advice and guidance.
  ZG: calibrated redshift distributions from COSMOS 30-band data.
  EG: galaxy sample selection and shot noise estimation.
  RM: provided guidance on how to use the HSC data.
  HM: provided guidance on how to use the HSC data.
  JN: assistance in interpretation of galaxy clustering results.
  SS: model parametrization, testing MCMC chains, studied uncertainties in the first bin's $N(z)$.
  EW: produced observing condition maps.
